---
title: "PSW_Cleaning_For_Merging"
output: html_document
date: "2024-10-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal of this R markdown file is to clean the PSW data into a format that allows merging with the random forests weather station data

Load the libraries
```{r}
library(raster)
library(sf)
library(doParallel)
library(tidyr)
```

Let's set up the shapefile to cast the PSW data correctly
```{r}
us_shp <- read_sf("/Volumes/Seagate Portable Drive/Shapefiles/tl_2022_us_state/tl_2022_us_state.shp")

#Subset on the Connecticut shapefile
ct_shp <- us_shp[us_shp$NAME == "Connecticut",]

#Get the lon-lat crs
lon_lat_crs <- st_crs(ct_shp)

#Get the CRS of the shapefile
crsSH <- CRS("+init=epsg:4269")
```


Now let's define the function for the loop we will be using
```{r}
#Filename is the name of the specific month_year combo we are processing
#var_name is the name of the variable (prcp_avg, srad_avg, etc.)
process_psw <- function(filename, var_name){
  #Create a notice of what we are processing
  print(paste0("Currently processing: ", filename))
  #Load the tif file, project the raster and mask the raster
  raster_data <- raster::brick(filename)
  raster_data <- raster::projectRaster(raster_data, crs = crsSH)
  raster_data <- raster::mask(raster_data, mask = ct_shp)
  
  #Now we want to extract the values and coordinates to bind the two data variables together
  raster_values <- data.frame(raster::values(raster_data))
  raster_coords <- data.frame(raster::coordinates(raster_data))
  
  #Now we must do the following to the values:
    #1) Turn the column names into actual dates using the filename which is i
    #2) Apply the correct transformation on the values to scale them. Seems like for the ERA5 reanalysis data, we do not need to transform the values
    #3) Cbind the result with raster_coords
    #4) Pivot the dates 
  
  #Let's get the dates
  month_year <- regmatches(filename, regexec("_([0-9]{2})_([0-9]{4})", filename))
  month <- as.integer(month_year[[1]][2])
  year <- as.integer(month_year[[1]][3])
  
  #Now we will get the sequence of dates for the columns. Note, we can actually get to the end date by adding the number of columns we have minus 1
  #in our raster_values
  start_date <- as.Date(paste(year, month, "01", sep = "-"))
  end_date <- as.Date(start_date, format = "%Y-%m-%d") + ncol(raster_values) - 1
  date_seq <- seq(start_date, end_date, by = "day")
  
  #Great, now we can use these to rename the raster_values columns
  names(raster_values) <- date_seq
  
  #We have some negative values in the data and that doesn't make sense, so let's get rid of all the negative values and turn them into NAs once we pivot them
  raster_values <- cbind(raster_coords, raster_values)
  
  #Now we pivot, note how we will need to supply the "values_to" argument later on
  raster_values <- tidyr::pivot_longer(raster_values, cols = 3:ncol(raster_values), names_to = "Date", values_to = var_name)
  
  #And turn the dates into a date format
  raster_values$Date <- as.Date(raster_values$Date)
  
  summary(raster_values[, 4])
  
  #And then deal with the negative values
  if(!(var_name %in% c("u_wind", "v_wind"))){
    raster_values[raster_values[, 4] < 0 & !is.na(raster_values[, 4]), 4] <- NA
  }
  
  #Return the data frame
  return(raster_values)
}
```

Now let's create the loop using the basepath
```{r}
var_dir <- "/Volumes/Seagate Portable Drive/0.raw_data/0.8.PrecipitationSolarWind/u_wind"

#Now we get the avg, min, and max files
avg_files <- list.files(path = list.files(path = var_dir, pattern = "avg", full.names = TRUE), full.names = TRUE)
min_files <- list.files(path = list.files(path = var_dir, pattern = "min", full.names = TRUE), full.names = TRUE)
max_files <- list.files(path = list.files(path = var_dir, pattern = "max", full.names = TRUE), full.names = TRUE)

#Combine them all into separate elements of a list
sum_files <- list(avg_files, min_files, max_files)

#Create a list of the years
data_years <- c("2019", "2020", "2021", "2022", "2023")

cl <- makeCluster(5)
registerDoParallel(cl)

#Go through the values
for(file_set in sum_files){
  #Grab the var name
  var_name <- gsub("_(\\d{2}_\\d{4})\\.grib", "", basename(file_set[1]))
  
  #We can create a loop using the years
  for(var_year in data_years){
    #Define the files
    agg_data <- grep(var_year, file_set, value = TRUE)
    
    #Now we create the data frame
    data <- data.frame()
    
    #Run the for each loop
    data <- foreach(filename = agg_data, .combine = "rbind") %dopar% {
      process_psw(filename, var_name)
    }
    
    #Now we need to save the file in the form of an RDS file
    clean_file_path <- "/Volumes/Seagate Portable Drive/3.clean_data/Cleaned_Data/"
    
    #Create the file path
    save_file_path <- paste0(clean_file_path, "ERA_", basename(var_dir), "/", var_name, "_", var_year, ".rds")
    
    #Check if the directory exists and create it if not
    dir.create(dirname(save_file_path), recursive = TRUE, showWarnings = FALSE)
    
    #Now we save the file
    saveRDS(data, file = save_file_path)
  }
}

stopCluster(cl)
```


Test the values
```{r}
test <- readRDS("/Volumes/Seagate Portable Drive/3.clean_data/Cleaned_Data/ERA_v_wind/vwind_max_2019.rds")
test_2 <- readRDS("/Volumes/Seagate Portable Drive/3.clean_data/Cleaned_Data/ERA_v_wind/vwind_avg_2019.rds")
test_3 <- readRDS("/Volumes/Seagate Portable Drive/3.clean_data/Cleaned_Data/ERA_v_wind/vwind_min_2019.rds")

summary(test$vwind_max)
summary(test_2$vwind_avg)
summary(test_3$vwind_min)

test

unique(test$x)
```

_________________________________-

Let's define the directory. We will go one weather variable at a time
```{r}
#The directory of the weather variable
var_dir <- "/Volumes/Seagate Portable Drive/0.raw_data/0.8.PrecipitationSolarWind/precipitation"

#Now we can get the avg, min, and max files
avg_files <- list.files(path = list.files(path = var_dir, pattern = "avg", full.names = TRUE), full.names = TRUE)
min_files <- list.files(path = list.files(path = var_dir, pattern = "min", full.names = TRUE), full.names = TRUE)
max_files <- list.files(path = list.files(path = var_dir, pattern = "max", full.names = TRUE), full.names = TRUE)
```

Create separate lists of file names to get the dataframes grouped by year
```{r}
avg_2019 <- grep("2019", avg_files, value = TRUE)
avg_2020 <- grep("2020", avg_files, value = TRUE)
avg_2021 <- grep("2021", avg_files, value = TRUE)
avg_2022 <- grep("2022", avg_files, value = TRUE)
avg_2023 <- grep("2023", avg_files, value = TRUE)
```


Let's create the skeleton of the function. We note how total precipitation is in meters.
```{r}
data <- data.frame()
for(i in avg_2019[1:3]){
  
  #Create a notice of what we are processing
  print(paste0("Currently processing: ", i))
  #Load the tif file, project the raster and mask the raster
  raster_data <- raster::brick(i)
  raster_data <- raster::projectRaster(raster_data, crs = crsSH)
  raster_data <- raster::mask(raster_data, mask = ct_shp)
  
  #Now we want to extract the values and coordinates to bind the two data variables together
  raster_values <- data.frame(raster::values(raster_data))
  raster_coords <- data.frame(raster::coordinates(raster_data))
  
  #Now we must do the following to the values:
    #1) Turn the column names into actual dates using the filename which is i
    #2) Apply the correct transformation on the values to scale them. Seems like for the ERA5 reanalysis data, we do not need to transform the values
    #3) Cbind the result with raster_coords
    #4) Pivot the dates 
  
  #Let's get the dates
  month_year <- regmatches(i, regexec("_([0-9]{2})_([0-9]{4})", i))
  month <- as.integer(month_year[[1]][2])
  year <- as.integer(month_year[[1]][3])
  
  #Now we will get the sequence of dates for the columns. Note, we can actually get to the end date by adding the number of columns we have minus 1
  #in our raster_values
  start_date <- as.Date(paste(year, month, "01", sep = "-"))
  end_date <- as.Date(start_date, format = "%Y-%m-%d") + ncol(raster_values) - 1
  date_seq <- seq(start_date, end_date, by = "day")
  
  #Great, now we can use these to rename the raster_values columns
  names(raster_values) <- date_seq
  
  #We have some negative values in the data and that doesn't make sense, so let's get rid of all the negative values and turn them into NAs once we pivot them
  raster_values <- cbind(raster_coords, raster_values)
  
  #Now we pivot, note how we will need to supply the "values_to" argument later on
  raster_values <- tidyr::pivot_longer(raster_values, cols = 3:ncol(raster_values), names_to = "Date", values_to = "prcp_avg")
  
  #And turn the dates into a date format
  raster_values$Date <- as.Date(raster_values$Date)
  
  #And then deal with the negative values
  raster_values[raster_values[, 4] < 0 & !is.na(raster_values[, 4]), 4] <- NA
  
  #Now we rbind
  data <- rbind(data, raster_values)
}

```













